---
title: "Descriptives"
author: "Marc Comas-Cufí, Antonella Buccianti, Josep Antoni Martín-Fernandez and Glòria Mateu-Figueras"
date: "18 Oct 2015"
output: 
  html_document: 
    toc: yes
---

```{r, include=F}
library(dplyr)
library(ggplot2)
library(tidyr)
library(compositions)
library(ggmap)
knitr::opts_chunk$set(comment = "  ")
options(dplyr.print_min = 5, dplyr.print_max = 30, width=100)
load('data/geowater.RData')
load('data/geowater_maps.RData')
source('functions.R')
```

# Location

Locations at no more than 15 meters are considered one unique location.

```{r}
# By considering only 4 decimal digits we are assuming 15 meters between observations
data = df.coord %>% 
  mutate('loc.id' = sprintf('%.4f,%.4f', lat, lon) ) %>%
  left_join(df.date, by = 'id') %>%
  left_join(df.chem, by = 'id') %>%
  mutate(loc.id = ifelse(loc.id == 'NA,NA', NA, loc.id))
```

## Non-registered measurements

The location was no registered in `r sum(is.na(data['loc.id']))` observations.

```{r}
data['loc.id'] %>% 
  is.na %>% 
  ifelse('Non-registered', 'Registered') %>%
  table %>% 
  addmargins
```

## Measurements at the same location

There were `r sum(!is.na(data$loc.id))` different observations with registered locatio and there were `r length(table(data$loc.id))` different locations.

```{r}
data %>% filter(!is.na(loc.id)) %>% {table(.$loc.id)} %>% table %>% addmargins
```

```{r, echo=F, eval=F}
ggplot() + 
  geom_bar(data = data %>% filter(!is.na(loc.id)) %>% 
             group_by(loc.id) %>% summarise(nloc = n()) %>%
             group_by(nloc) %>% summarise(n = n()),
  aes(x = nloc, y = n), stat = 'identity') + theme_classic()
```

```{r}
with(data %>% filter(!is.na(loc.id)) %>% count(loc.id), 
     barplot(table(n), main='Repeated measures', xlab = 'Number of repetitions'))
```

## Total variance in each location

```{r}
totalVariance = function(.data) .data %>% compo %>% acomp %>% variation %>% 
{data.frame('totalVariance' = sum(.))}
```

For all the observations in each sampling point the total variance was calculated.

```{r}
(tv.sample <- data %>% group_by(loc.id) %>%
  do(totalVariance(.)) %>%
   left_join(data %>% count(loc.id), by='loc.id') %>% filter(!is.na(totalVariance)) %>% ungroup %>% arrange(-totalVariance) )
```

```{r, collapse=TRUE, results='hold'}
opar = par(mfrow=c(1,2),oma=c(0,0,2,0))
boxplot(sqrt(tv.sample$totalVariance), main=NULL)
abline(h=sqrt(tv.sample$totalVariance[is.na(tv.sample$loc.id)]), col=2)

hist(sqrt(tv.sample$totalVariance), main=NULL, xlab='Total var. (square root)')
abline(v=sqrt(tv.sample$totalVariance[is.na(tv.sample$loc.id)]), col=2)
title("Total variance distribution for sampling points", outer=TRUE)
```

```{r, results='asis', echo=F}
LOC.ID = '43.4743,11.8952'
lat = strsplit(LOC.ID, ',')[[1]][1]
lon = strsplit(LOC.ID, ',')[[1]][2]
cat(sprintf('<p>The location <a href="http://maps.google.com/maps?q=%s,%s">(lat:%s-lon:%s)</a> is the location with higher total variance.</p>', 
        lat, lon, lat, lon))
```

```{r, results='hold'}
data %>% subset(loc.id == LOC.ID) %>% select(loc.id, year, month, ph, conduct, HCO3:SO4)
data %>% subset(loc.id == LOC.ID) %>% compo %>% closure
```

 * Other locations (for each location compositiona and its closure is shown):

```{r, results='asis', echo=F}
LOC.ID = '43.6582,10.5785'
lat = strsplit(LOC.ID, ',')[[1]][1]
lon = strsplit(LOC.ID, ',')[[1]][2]
cat(sprintf('<p>Location <a href="http://maps.google.com/maps?q=%s,%s">(lat:%s-lon:%s)</a>.</p>', 
        lat, lon, lat, lon))
```

```{r, results='hold'}
data %>% subset(loc.id == LOC.ID) %>% select(loc.id, year, month, ph, conduct, HCO3:SO4)
data %>% subset(loc.id == LOC.ID) %>% compo %>% closure
```

```{r, results='asis', echo=F}
LOC.ID = '43.4774,11.8932'
lat = strsplit(LOC.ID, ',')[[1]][1]
lon = strsplit(LOC.ID, ',')[[1]][2]
cat(sprintf('<p>Location <a href="http://maps.google.com/maps?q=%s,%s">(lat:%s-lon:%s)</a>.</p>', 
        lat, lon, lat, lon))
```

```{r, results='hold'}
data %>% subset(loc.id == LOC.ID) %>% select(loc.id, year, month, ph, conduct, HCO3:SO4)
data %>% subset(loc.id == LOC.ID) %>% compo %>% closure
```

```{r, results='asis', echo=F}
LOC.ID = '43.7109,10.7837'
lat = strsplit(LOC.ID, ',')[[1]][1]
lon = strsplit(LOC.ID, ',')[[1]][2]
cat(sprintf('<p>Location <a href="http://maps.google.com/maps?q=%s,%s">(lat:%s-lon:%s)</a>.</p>', 
        lat, lon, lat, lon))
```

```{r, results='hold'}
data %>% subset(loc.id == LOC.ID) %>% select(loc.id, year, month, ph, conduct, HCO3:SO4)
data %>% subset(loc.id == LOC.ID) %>% compo %>% closure
```


## Sample locations

The locations are plot in quartiles with respect the total variance. The colors are shown with respect the logarithm of this total variance.

```{r, fig.width=8, warning=F}
xquantile = function(X, K){
  cut(X, quantile(X, probs = seq(0, 1, length.out = K+1)), include.lowest = T)
}
df = data_frame(
  lat = as.numeric(sapply(strsplit(tv.sample[['loc.id']], ","), function(v) v[1])),
  lon = as.numeric(sapply(strsplit(tv.sample[['loc.id']], ","), function(v) v[2])),
  totalVariance = tv.sample[['totalVariance']] ) %>% 
  filter(totalVariance > 0.01 & !is.na(totalVariance)) %>%
  mutate(
    quartiles = xquantile(totalVariance, 4))

ggmap(map_bw) + 
  geom_point(data = df %>% filter(totalVariance > 0), aes(x=lon, y=lat, col=log(totalVariance)), size=2) +
  facet_wrap(~quartiles, nrow=2) +
  scale_colour_gradient(high="red", low='blue')
```

# Temporal data

## Year distribution


```{r, results='hold'}
ifelse(!is.na(data[['year']]), 'available', 'non-available') %>% table %>% addmargins
data[['year']] %>% table(useNA = 'ifany')
```

## Month distribution

```{r, results='hold'}
ifelse(!is.na(data[['month']]), 'available', 'non-available') %>% table %>% addmargins
data[['month']] %>% table(useNA = 'ifany')
```

 * The total Variance

```{r}
(tv.month <- data %>% group_by(month) %>%
  do(totalVariance(.)) %>%
   left_join(data %>% count(month), by='month') %>% filter(!is.na(totalVariance)) %>% ungroup %>% arrange(-totalVariance) )
```

```{r, fig.width=9, warning=F}
ggmap(map_bw) + 
  geom_point(data = data %>% filter(!is.na(month)), aes(x=lon, y=lat), size=2) +
  facet_wrap(~month, nrow=2)
```

# Chemical composition

## Describing each component

```{r}
summarise_components = function(.data) .data %>%
  gather(key=component, value=val) %>%
  group_by(component) %>%
  summarise(
    na = sum(is.na(val)),
    min = min(val, na.rm=T),
    q1 = quantile(val, 0.25, na.rm=T),
    median = median(val, na.rm=T),
    q3 = quantile(val, 0.75, na.rm=T),
    max = max(val, na.rm=T) )
```

 * Raw data

```{r}
data %>% 
  select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% 
  summarise_components
```

* Closured data (100%)

```{r}
data %>% 
  select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% closure %>% 
  summarise_components
```

* Closured data by month

```{r}
df.res = data %>% 
  select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% closure %>% 
  mutate('month' = data[['month']]) %>%
  gather(key=component, value=val, -month) %>%
  group_by(month, component) %>%
  summarise(
    na = sum(is.na(val)),
    min = min(val, na.rm=T),
    q1 = quantile(val, 0.25, na.rm=T),
    median = median(val, na.rm=T),
    q3 = quantile(val, 0.75, na.rm=T),
    max = max(val, na.rm=T) )
```

```{r, fig.with = 9}
ggplot() +
  geom_point(data = df.res, aes(x = as.factor(month), y = median)) +
  geom_errorbar(data = df.res, aes(x = as.factor(month), ymin = q1, ymax = q3)) +
  facet_wrap(~component, scales = 'free', nrow = 2) + xlab('Month') + ylab('%')
```


