---
title: "Descriptives"
author: "Marc Comas-Cufí, Antonella Buccianti, Josep Antoni Martín-Fernandez and Glòria Mateu-Figueras"
date: "18 Oct 2015"
output: 
  html_document: 
    toc: yes
---

```{r, include=F}
library(dplyr)
library(ggplot2)
library(tidyr)
library(compositions)
options(dplyr.print_min = 5, dplyr.print_max = 30, width=100)
load('data/geowater.RData')

compo = function(.data) .data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% subset(apply(., 1, prod) != 0)
closure = function(.data, k = 100) (k*.data/apply(.data, 1, sum)) %>% round(4) %>% tbl_df

gmean = function(...){
  .data = data.frame(list(...))
  apply(.data, 1, prod)^(1/NCOL(.data))
}

variation_array = function(.data, output='m.v'){
  vrs = names(.data)
  bind_cols(.data %>% setNames(sprintf('col.%s', names(.))),
            .data %>% setNames(sprintf('row.%s', names(.)))) %>%
    gather(key=row.var, value=row.val, row.HCO3:row.SO4) %>%
    gather(key=col.var, value=col.val, col.HCO3:col.SO4) %>%
    group_by(col.var, row.var) %>% summarize(
  'm' = mean(log(row.val/col.val)),
  'v' = var(log(row.val/col.val)),
  'm.v' = sprintf("  %3.1f (%3.1f)", m, v)) %>% 
    select_('row.var', 'col.var', .dots=output) %>%
    spread_('col.var', output)
}


balance = function(.data) .data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>%
  mutate(
    b1 = sqrt(4*3/(4+3)) * log(gmean(Ca, Mg, Na, K) / gmean(HCO3, SO4, Cl)),
    b2 = sqrt(2*2/(2+2)) * log(gmean(Ca, Mg) / gmean(Na, K)),
    b3 = sqrt(1*1/(1+1)) * log(Ca/Mg),
    b4 = sqrt(1*1/(1+1)) * log(Na/K),
    b5 = sqrt(1*2/(1+2)) * log(gmean(HCO3)/gmean(SO4, Cl)),
    b6 = sqrt(1*1/(1+1)) * log(SO4/Cl) ) %>% select(b1,b2,b3,b4,b5,b6)
```

# Location

Locations at no more than 15 meters are considered one unique location.

```{r}
# By considering only 4 decimal digits we are assuming 15 meters between observations
data = df.coord %>% 
  mutate('loc.id' = sprintf('%.4f-%.4f', lat, lon) ) %>%
  left_join(df.date, by = 'id') %>%
  left_join(df.chem, by = 'id') %>%
  mutate(loc.id = ifelse(loc.id == 'NA-NA', NA, loc.id))
```

## Non-registered measurements

The location was no registered in `r sum(is.na(data['loc.id']))` observations.

```{r}
data['loc.id'] %>% 
  is.na %>% 
  ifelse('Non-registered', 'Registered') %>%
  table %>% 
  addmargins
```

## Measurements at the same location

```{r, echo=F, eval=F}
ggplot() + 
  geom_bar(data = data %>% filter(!is.na(loc.id)) %>% 
             group_by(loc.id) %>% summarise(nloc = n()) %>%
             group_by(nloc) %>% summarise(n = n()),
  aes(x = nloc, y = n), stat = 'identity') + theme_classic()
```

```{r}
with(data %>% filter(!is.na(loc.id)) %>% count(loc.id), 
     barplot(table(n), main='Repeated measures', xlab = 'Number of repetitions'))
```

## Total variance in each location

```{r}
totalVariance = function(.data) .data %>% compo %>% acomp %>% variation %>% 
  {data.frame('totalVariance' = sum(.))}
```

For all the observations in each sampling point the total variance was calculated.

```{r}
(tv.sample <- data %>% group_by(loc.id) %>%
  do(totalVariance(.)) %>% filter(!is.na(totalVariance)) %>% ungroup %>% arrange(-totalVariance) )
```

```{r, collapse=TRUE, results='hold'}
opar = par(mfrow=c(1,2),oma=c(0,0,2,0))
boxplot(sqrt(tv.sample$totalVariance), main=NULL)
abline(h=sqrt(tv.sample$totalVariance[is.na(tv.sample$loc.id)]), col=2)

hist(sqrt(tv.sample$totalVariance), main=NULL, xlab='Total var. (square root)')
abline(v=sqrt(tv.sample$totalVariance[is.na(tv.sample$loc.id)]), col=2)
title("Total variance distribution for sampling points", outer=TRUE)
```

```{r, results='asis', echo=F}
LOC.ID = '43.4743-11.8952'
lat = strsplit(LOC.ID, '-')[[1]][1]
lon = strsplit(LOC.ID, '-')[[1]][2]
cat(sprintf('<p>The location <a href="http://maps.google.com/maps?q=%s,%s">(lat:%s-lon:%s)</a> is the location with higher total variance</p>', 
        lat, lon, lat, lon))
```

```{r}
data %>% subset(loc.id == LOC.ID) %>% select(loc.id, year, month, ph, conduct, HCO3:SO4)
data %>% subset(loc.id == LOC.ID) %>% compo %>% closure
```

# Chemical composition

## Describing each component

```{r}
summarise_components = function(.data) .data %>%
  gather(key=component, value=val) %>%
  group_by(component) %>%
  summarise(
    na = sum(is.na(val)),
    min = min(val, na.rm=T),
    q1 = quantile(val, 0.25, na.rm=T),
    median = median(val, na.rm=T),
    q3 = quantile(val, 0.75, na.rm=T),
    max = max(val, na.rm=T) )
```

 * Raw data

```{r}
data %>% 
  select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% 
  summarise_components
```

* Closured data (100%)

```{r}
data %>% 
  select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% closure %>% 
  summarise_components
```


